How environmental drivers of synchrony interact

### 2021-12-22 Paper/Paper.Rnw
small edits while reading the paper to get back up to speed

%header material for paper
\title{How environmental drivers of spatial synchrony interact}
\date{}

%Alternative titles
%Interactions among Moran effects
%Interacting Moran effects
%How Moran effects interact
%On interactions between Moran effects
%Interactions between environmental drivers of spatial synchrony
%Environmental drivers of spatial synchony can strongly interact
%Environmental drivers of spatial synchrony interact

@@

%Tentative broad outline:
%DONE 1) Intro spectral quantities, the spectrum, the cospectrum, and the cross spectrum of two time series/processes. And the spectral
%matrix for putting them all in one structure.
%DONE 2) Define your measures of timescale/frequency-specific synchrony and crosss-variable synchrony, explain why these make sense as
%ways to measure those things.
%DONE 3) Present the model and the theory
%4) Present "toy" examples and fig 1 (need to think a lot about these). DAN: perhaps these actually belong in Methods (describe the specific-case models we will consider) and results (apply the general theory to those specific cases)
%
%Notes:
%Compared to the deer paper, this is a much deeper presentation of the theory in the main text. I'm not sure that's avoidable.
%For one thing, the theory is explicitly applied to get the results, later, and that was not the case for the deer paper.

@@

\noindent We now describe how we pursue goal G2 of the Introduction, to develop 
theoretical case studies that illuminate the intuition of interacting Moran effects. 
We develop case studies A-C. For all cases, the model time step was assumed to be one quarter, i.e., 
four model time steps per year. This makes no real mathematical difference, but was done to 
facilitate later comparisons with results based on kelp data, which were available on a
quarterly sampling frequency. 
For case study A,  the environmental variable 
$\epsilon^{(1)}$ is assumed to have a simple positive effect on populations, but 
lagged by one model time step, i.e.,
$p_0^{(1)}=0$, $p_1^{(1)}>0$, and $p_i^{(1)}=0$ for $i>1$. 
For case study B, which generalizes A, $\epsilon^{(1)}$ is again assumed to have a simple 
positive effect on populations,
but now lagged by $n$ time steps, so that $p_n^{(1)}>0$ while $p_i^{(1)}=0$ for $i \neq n$.
For both A and B, the effects of $\epsilon^{(2)}$
are assumed to be un-lagged, i.e., $p_0^{(2)} \neq 0$ and $p_i^{(2)}=0$ for $i>0$.
For case study C, the effects of $\epsilon^{(1)}$ are lagged by $n_1$ and the 
effects of $\epsilon^{(2)}$ are lagged by $n_2$ time steps, i.e.,
$p_{n_1}^{(1)}>0$ while $p_i^{(1)}=0$ for $i \neq n_1$, and 
$p_{n_2}^{(2)}>0$ while $p_i^{(2)}=0$ for $i \neq n_2$.
The noise process
$(\epsilon^{(1)},\epsilon^{(2)})$ is assumed to be a Gaussian white-noise process 
for both A and B, i.e., 
the random variables $(\epsilon^{(1)}(t),\epsilon^{(2)}(t))$ are independent and 
identically distributed (iid) for distinct times, $t$. Noise may be correlated across space,
and the components of $\epsilon^{(1)}(t)$ may be correlated with those of $\epsilon^{(2)}(t)$.
For C, the noise processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are
each assumed to exhibit spatially synchronous periodic oscillations of period equal to one year, i.e.,
four model time steps, but with different phases. Peaks in $\epsilon^{(1)}$ are assumed to 
either lead or lag (we consider two model scenarios) peaks in $\epsilon^{(2)}$ by one quarter.
Such a scenario could be realized by annually 
periodic environmental fluctuations sampled quarterly,
e.g., kelp wave action in central California peaks annually in the winter, whereas surface-water
nitrate concentrations also fluctuate with period one year, but peak in the spring, a delay of one quarter
compared to the wave peak. 
Details of the noise processes are specified in SI section \ref{SM-sect:examples_cross_sync}.
For all case studies,
the autoregressive order of population dynamics is assumed, for simplicity, to be $1$, i.e, $b_1 \neq 0$ 
but $b_i=0$ for $i>1$.

%Tentative broad outline
%1) Para from Tom on how remote sensing data are gathered and coverted into what he handed to me. Brief, mostly 
%citing other works where it is laid out in much more detail. 
%2) Para from Tom on how he got to the wave and NO3 data
%3) Para summarizing the data cleaning I did
%Comment) Actually, might want to send the abive 3 paras to sup mat, and just have one para describing the final format of the
%data, especially if you make this paper also be about plankton. But do that later if you do it.
%4) Some paras about the plankton data, similar structure to how the kelp data were presented (and using sup mat as described)
%5) Para about the linear modelling - this is how certain parts of the main theoretical equation were estimated
%6) Para about spectral estimates. Brief, cites appropriate parts of Brillinger, uses sup mat if needed. This is how certain
%other parts of the main theoretical equation were estimated. 
%Comment) I could write these now, but then if I end up having to do something slightly different for plankton, I'll have to 
%rewrite them. Maybe now is the time to decide if I want this to be a kelp-only or a kelp-an-plankton paper and then do the
%plankton work.
%Comment) Somehow need to get across what was done - different parts of main equation were examined to explore the different
%aspects of how the mechanism works for the focal taxa. Elaborate as appropriate.

@@

%Para basically confirming earlier work, as a starter
Our theoretical results demonstrate that interaction effects between Moran drivers: 1) can be 
comparable in strength to direct Moran effects; 2) can be either synergistic or destructive, i.e., positive or negative; and 3) 
can depend strongly on timescale, even having different signs on different timescales 
(Fig. \ref{fig:TheoryCaseA}a,e,i,m). First, for all case studies, the magnitude of 
interaction effects was comparable to that of direct Moran effects (compare the dashed and solid lines on 
Fig. \ref{fig:TheoryCaseA}a,e,i,m). So interactions can contribute substantially to overall population 
synchrony. Second, in contrast to the direct Moran effects of a single environmental variable, which provide non-negative 
contributions to population synchrony at all timescales for which environmental synchrony 
is positive, interactions between Moran effects can be positive or negative. For instance, case study A
showed negative interaction effects at short timescales (Fig. \ref{fig:TheoryCaseA}a); and, for some
parameters, case study C showed negative interactions at annual timescales (Fig. \ref{fig:TheoryCaseA}m).
Thus interaction effects can either augment or reduce population synchrony from what its value would be 
if only direct Moran effects operated. Finally, interactions between Moran effects depended strongly on timescale 
for all case studies (Fig. \ref{fig:TheoryCaseA}a,e,i,m), and differed in sign across timescales for case studies
A and B and for one of the C scenarios. This augments earlier studies demonstrating that direct Moran effects can 
depend on timescale \citep{Defriez2016,Sheppard2016,Desharnais2018,Anderson2021}. The points made in this paragraph
also follow from the results of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp}, though not necessarily
straightforwardly. Additionally, those are case studies, so our results also provide novel theoretical generality;
and our theory can be interpreted to provide intuition about how interacting Moran effects work,
as follows.

@@

Case study B reveals how Moran effects can interact when lags are longer than 1 model time step. 
For B, $\epsilon^{(1)}$ effects on populations were lagged by 3 quarters but 
$\epsilon^{(2)}$ effects were unlagged (Methods). Thus interactions between Moran effects were 
negative on the shortest timescales 
(0.5 years = 2 quarters; Fig. \ref{fig:TheoryCaseA}e, left side of panel). This is because the 
population influences of Moran effects of timescale 2
quarters and lagged by 3 quarters are exactly out of phase with unlagged Moran effects of timescale
2 quarters, producing negative interactions. 
Interaction effects on long timescales were instead synergistic, i.e., positive, for case study B 
(Fig. \ref{fig:TheoryCaseA}e, right side of panel), for the same reason as case study A (see above).
At timescale 3 quarters, Moran effects are, of course, reinforcing because the lag equals the timescale,
producing Moran effects of the two drivers which are in phase 
(Fig. \ref{fig:TheoryCaseA}e, peak at timescale 3 quarters = 0.75 years). 
At a timescale of 1 year, a lag of 3 quarters causes the Moran effects of 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ to be $3/4$-phase out of alignment, resulting in zero interaction effects. 
Likewise at a timescale of 3 years (12 quarters), a lag of 3 quarters 
makes the Moran effects of 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ be $1/4$-phase out of alignment, again producing zero interaction. 
Finally, at timescale 1.5 years (6 quarters), the 3-quarter lag equals half the timescale, so 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ effects counteract each other, producing negative 
interactions. Thus interactions between Moran drivers were zero 
at timescales 1 and 3 years for B, were 
negative for short timescales and for timescale 1.5 years, and were positive for 
timescale 0.75 years and for long timescales (Fig. \ref{fig:TheoryCaseA}e).
Theory captures these complexities because 
the size of the 3-quarter phase lag of this scenario, relative to a given timescale,
is captured by $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}f).
Because $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number for case study B,
the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
is the same as that of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}f); 
that common phase then controls the sign of 
$2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, which is the 
interaction effect term in (\ref{rhoww_T1to3}).

Cases A and B assumed white noise processes and real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$,
but case study C illustrates what can happen when noise processes have periodic components, e.g., 
both noise processes oscillate with annual periodicity according to the seasons, an important scenario, ecologically. 
What matters then is whether and how the lag between the two seasonal noise processes aligns with the difference 
between the lags of their effects on populations, as we now explain. 
For the scenarios we considered under case study C (Fig. \ref{fig:TheoryCaseA}i-p), 
$\epsilon^{(1)}$ effects on populations were lagged (i.e., delayed in their influence) by 1 quarter, but 
$\epsilon^{(2)}$ effects were immediate (Methods and Fig. \ref{fig:TheoryCaseA} caption).
For scenario 1 of case study C (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
noise process $\epsilon^{(2)}$ lagged peaks in the periodic process $\epsilon^{(1)}$ by
1 quarter. Thus lags cancelled and the annual peak population effects of $\epsilon^{(1)}$ and
$\epsilon^{(2)}$ arrived in sync with each other each year, producing positive interactions 
between Moran effects on annual timescales (Fig. \ref{fig:TheoryCaseA}i), and more population
synchrony than would have been expected if Moran effects operated independently. Contrastingly,
for scenario 2 of case study C (Fig. \ref{fig:TheoryCaseA}m-p), peaks in the periodic
noise process $\epsilon^{(1)}$ lagged peaks in the periodic process $\epsilon^{(2)}$ by
1 quarter. Thus lags compounded and the annual peak population effects of $\epsilon^{(1)}$
arrived 2 quarters after those of $\epsilon^{(2)}$ every year, a half-phase shift for an 
annually oscillating effect, which corresponds to negative interactions between Moran influences
and less population synchrony than would have been expected if Moran effects operated independently 
(Fig. \ref{fig:TheoryCaseA}m). The only difference between scenarios 1 and 2 of case study C
is the relative timing of annual peaks in $\epsilon^{(1)}$ and $\epsilon^{(2)}$.
Our theory captures these complexities because $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$
quantifies the relative phase lag between the effects of $\epsilon^{(1)}$ and 
$\epsilon^{(2)}$ relative to a given timescale, while 
$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ captures the phase lag between 
oscillations in the variables themselves, relative to the timescale.
The phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
therefore captures whether relative lags of effects and lags between 
the variables themselves align or misalign, with the degree of alignment
reflected in the real part of this quantity.

@@

\begin{figure}
\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseA.pdf}\\
\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseB.pdf}\\
\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v01.pdf}\\
\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v02.pdf}
\caption{Theoretical case studies A (panels a-d), B (panels e-h) and C (i-p, two alternative sets of 
parameters used for i-l and m-p, respectively). Left panels (a, e, i, m) show the terms on the right side
of (\ref{rhoww_T1to3}). On those panels, the green line is 
$\frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(1)} \epsilon^{(1)}}$, quantifying 
population synchrony due to the direct Moran effects of $\epsilon^{(1)}$. The blue line
is $\frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(2)} \epsilon^{(2)}}$, quantifying 
population synchrony due to the direct Moran effects of $\epsilon^{(2)}$.
The dashed green-blue line is 
$\frac{2 \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right)}{|f_{\mcB}|^2}$, 
which is population synchrony due to interacting Moran effects.
%***DAN: More text here as we flesh out the presentation - need to describe the middle two columns of panels 
The functions plotted on d, h, l, p are those on a, e, i, m, respectively, times 
$|f_{\mcB}|^2$, thereby showing the influence of autoregressive population effects on
synchrony [see (\ref{rhoww_T1to3})-(\ref{rhoww_OC})].
Model parameters were: $N=5$ and $b_1=0.4$ for all case studies; 
$p_1^{(1)}=1.1$ and $p_0^{(2)}=0.8$ for case study A; 
$p_3^{(1)}=1.1$ and $p_0^{(2)}=0.8$ for case study B;
$p_1^{(1)}=3$, $p_0^{(2)}=1.5$ 
%$c_1=0$ and $c_2=-4/9$ 
for case study C. For C, peaks in the periodic noise process $\epsilon^{(2)}$
lagged peaks in the periodic process $\epsilon^{(1)}$ by 1 quarter
in panels i-l, and $\epsilon^{(1)}$ lagged $\epsilon^{(2)}$ by the same
amount in m-p.
See SI section \ref{SM-sect:examples_cross_sync} and Fig. \ref{SM-fig:rhosForC} for 
parameters associated with the noise for each case study,
and see Methods and SI section \ref{SM-sect:details_case_studies} for additional details.}\label{fig:TheoryCaseA}
\end{figure}

>>>>

%header material for paper
\title{How environmental drivers of synchrony interact}
\date{}

%Alternative titles
%Interactions among Moran effects
%Interacting Moran effects
%How Moran effects interact
%On interactions between Moran effects
%Interactions between environmental drivers of spatial synchrony
%Environmental drivers of spatial synchony can strongly interact
%Environmental drivers of spatial synchrony interact
%How environmental drivers of spatial synchrony interact

@@

@@

\noindent We now describe how we pursue goal G2 of the Introduction, to develop 
theoretical case studies that illuminate the intuition of interacting Moran effects. 
We develop case studies A-C. For all cases, the model time step was assumed to be one quarter, i.e., 
four model time steps per year. This makes no real mathematical difference, but was done to 
facilitate later comparisons with results based on kelp data, which were available with a
quarterly sampling frequency. 
For case study A,  the environmental variable 
$\epsilon^{(1)}$ is assumed to have a simple positive effect on populations, but 
lagged by one model time step, i.e.,
$p_0^{(1)}=0$, $p_1^{(1)}>0$, and $p_i^{(1)}=0$ for $i>1$. 
For case study B, which generalizes A, $\epsilon^{(1)}$ is again assumed to have a simple 
positive effect on populations,
but now lagged by $n$ time steps, so that $p_n^{(1)}>0$ while $p_i^{(1)}=0$ for $i \neq n$.
For both A and B, the effects of $\epsilon^{(2)}$
are assumed to be un-lagged, i.e., $p_0^{(2)} \neq 0$ and $p_i^{(2)}=0$ for $i>0$.
For case study C, the effects of $\epsilon^{(1)}$ are lagged by $n_1$ and the 
effects of $\epsilon^{(2)}$ are lagged by $n_2$ time steps, i.e.,
$p_{n_1}^{(1)}>0$ while $p_i^{(1)}=0$ for $i \neq n_1$, and 
$p_{n_2}^{(2)}>0$ while $p_i^{(2)}=0$ for $i \neq n_2$.
The noise process
$(\epsilon^{(1)},\epsilon^{(2)})$ is assumed to be a Gaussian white-noise process 
for both A and B. So 
the random variables $(\epsilon^{(1)}(t),\epsilon^{(2)}(t))$ are independent and 
identically distributed (iid) for distinct times, $t$. Noise may be correlated across space,
and the components of $\epsilon^{(1)}(t)$ may be correlated with those of $\epsilon^{(2)}(t)$.
For C, the noise processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are
each assumed to exhibit spatially synchronous periodic oscillations of period equal to one year, i.e.,
four model time steps, but with different phases. Peaks in $\epsilon^{(1)}$ are assumed to 
either lead or lag (we consider two model scenarios) peaks in $\epsilon^{(2)}$ by one quarter.
Such a scenario could be realized by annually 
periodic environmental fluctuations sampled quarterly,
e.g., kelp wave action in central California peaks annually in the winter, whereas surface-water
nitrate concentrations also fluctuate with period one year, but peak in the spring, a delay of one quarter
compared to the wave peak. 
Details of the noise processes are specified in SI section \ref{SM-sect:examples_cross_sync}.
For all case studies,
the autoregressive order of population dynamics is assumed, for simplicity, to be $1$, i.e, $b_1 \neq 0$ 
but $b_i=0$ for $i>1$.

%Tentative broad outline
%0) above stuff about the case studies 
%1) Para from Tom on how remote sensing data are gathered and coverted into what he handed to me. Brief, mostly 
%citing other works where it is laid out in much more detail. 
%2) Para from Tom on how he got to the wave and NO3 data
%3) Para summarizing the data cleaning I did
%Comment) Actually, might want to send the abive 3 paras to sup mat, and just have one para describing the final format of the
%data, especially if you make this paper also be about plankton. But do that later if you do it.
%5) Para about the linear modelling - this is how certain parts of the main theoretical equation were estimated
%6) Para about spectral estimates. Brief, cites appropriate parts of Brillinger, uses sup mat if needed. This is how certain
%other parts of the main theoretical equation were estimated. 
%Comment) I could write these now, but then if I end up having to do something slightly different for plankton, I'll have to 
%rewrite them. Maybe now is the time to decide if I want this to be a kelp-only or a kelp-an-plankton paper and then do the
%plankton work.
%Comment) Somehow need to get across what was done - different parts of main equation were examined to explore the different
%aspects of how the mechanism works for the focal taxa. Elaborate as appropriate.

@@

%Para basically confirming earlier work, as a starter
Our theoretical results demonstrate that interaction effects between Moran drivers: 1) can be 
comparable in strength to direct Moran effects; 2) can be either synergistic or destructive, i.e., positive or negative; and 3) 
can depend strongly on timescale, even having different signs on different timescales 
(Fig. \ref{fig:TheoryCaseA}a,e,i,m). First, for all case studies, the magnitude of 
interaction effects was comparable to that of direct Moran effects (compare the dashed and solid lines on 
Fig. \ref{fig:TheoryCaseA}a,e,i,m). So interactions can contribute substantially to overall population 
synchrony. Second, in contrast to the direct Moran effects of a single environmental variable, which provide non-negative 
contributions to population synchrony at all timescales for which environmental synchrony 
is positive, interactions between Moran effects can be positive or negative. For instance, case study A
showed negative interaction effects at short timescales (Fig. \ref{fig:TheoryCaseA}a); and, for some
parameters, case study C showed negative interactions at annual timescales (Fig. \ref{fig:TheoryCaseA}m).
Thus interaction effects can either augment or reduce population synchrony from what its value would be 
if only direct Moran effects operated. Finally, interactions between Moran effects depended strongly on timescale 
for all case studies (Fig. \ref{fig:TheoryCaseA}a,e,i,m), and differed in sign across timescales for case studies
A and B and for one of the C scenarios. This augments earlier studies demonstrating that direct Moran effects can 
depend on timescale \citep{Defriez2016,Sheppard2016,Desharnais2018,Anderson2021}. The points made in this paragraph
also follow from the results of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp}, though not necessarily
straightforwardly. Additionally, those are case studies, so our results provide novel theoretical generality;
and our theory can be interpreted to provide intuition about how interacting Moran effects work,
as follows.

@@

Case study B reveals how Moran effects can interact when lags are longer than 1 model time step. 
For B, $\epsilon^{(1)}$ effects on populations were lagged by $n$ quarters but 
$\epsilon^{(2)}$ effects were unlagged (Methods). We used $n=3$. Thus interactions between Moran effects were 
negative on the shortest timescales 
(0.5 years = 2 quarters; Fig. \ref{fig:TheoryCaseA}e, far left side of panel). This is because the 
population influences of Moran effects of timescale 2
quarters and lagged by 3 quarters are exactly out of phase with unlagged Moran effects of timescale
2 quarters, producing negative interactions. 
Interaction effects on long timescales were instead synergistic, i.e., positive, for case study B 
(Fig. \ref{fig:TheoryCaseA}e, right side of panel), for the same reason as case study A (see above).
At timescale 3 quarters, Moran effects are, of course, reinforcing because the lag equals the timescale,
producing Moran effects of the two drivers which are in phase 
(Fig. \ref{fig:TheoryCaseA}e, peak at timescale 3 quarters = 0.75 years). 
At a timescale of 1 year, a lag of 3 quarters causes the Moran effects of 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ to be $3/4$-phase out of alignment, resulting in zero interaction effects. 
Likewise at a timescale of 3 years (12 quarters), a lag of 3 quarters 
makes the Moran effects of 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ be $1/4$-phase out of alignment, again producing zero interaction. 
Finally, at timescale 1.5 years (6 quarters), the 3-quarter lag equals half the timescale, so 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ effects counteract each other, producing negative 
interactions. Thus interactions between Moran drivers were zero 
at timescales 1 and 3 years for B, were 
negative for short timescales and for timescale 1.5 years, and were positive for 
timescale 0.75 years and for long timescales (Fig. \ref{fig:TheoryCaseA}e).
Theory captures these complexities because 
the size of the 3-quarter phase lag of this scenario, relative to a given timescale,
is captured by $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}f).
Because $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number for case study B,
the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
is the same as that of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}g); 
that common phase then controls the sign of 
$2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, which is the 
interaction effect term in (\ref{rhoww_T1to3}).

Cases A and B assumed white noise processes and real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$,
but case study C illustrates what can happen when noise processes have periodic components, e.g., 
both noise processes oscillate with annual periodicity according to the seasons, an important scenario, ecologically. 
What matters then is whether and how the lag between the two seasonal noise processes aligns with the difference 
between the lags of their effects on populations, as we now explain. 
For the scenarios we considered under case study C (Fig. \ref{fig:TheoryCaseA}i-p), 
$\epsilon^{(1)}$ effects on populations were lagged (i.e., delayed in their influence) by 1 quarter, but 
$\epsilon^{(2)}$ effects were immediate (Methods and Fig. \ref{fig:TheoryCaseA} caption).
For scenario 1 of case study C (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
noise process $\epsilon^{(2)}$ lagged peaks in the periodic process $\epsilon^{(1)}$ by
1 quarter. Thus lags cancelled and the annual peak population effects of $\epsilon^{(1)}$ and
$\epsilon^{(2)}$ arrived in sync with each other each year, producing positive interactions 
between Moran effects on annual timescales (Fig. \ref{fig:TheoryCaseA}i), and more population
synchrony than would have been expected if Moran effects operated independently. Contrastingly,
for scenario 2 of case study C (Fig. \ref{fig:TheoryCaseA}m-p), peaks in the periodic
noise process $\epsilon^{(1)}$ lagged peaks in the periodic process $\epsilon^{(2)}$ by
1 quarter. Thus lags compounded and the annual peak population effects of $\epsilon^{(1)}$
arrived 2 quarters after those of $\epsilon^{(2)}$ every year, a half-phase shift for an 
annually oscillating effect, which corresponds to negative interactions between Moran influences
and less population synchrony than would have been expected if Moran effects operated independently 
(Fig. \ref{fig:TheoryCaseA}m). The only difference between scenarios 1 and 2 of case study C
is the relative timing of annual peaks in $\epsilon^{(1)}$ and $\epsilon^{(2)}$.
Our theory captures these complexities
because $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$
quantifies the relative phase lag between the effects of $\epsilon^{(1)}$ and 
$\epsilon^{(2)}$ relative to a given timescale, while 
$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ captures the phase lag between 
oscillations in the variables themselves, relative to the timescale (Fig. \ref{fig:TheoryCaseA}j,k,n,o).
The phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
therefore captures whether relative lags of effects and lags between 
the variables themselves align or misalign, with the degree of alignment
reflected in the real part of this quantity. On long timescales, interactions were the
same in the two secenaios, because on those timescales sub-annual lags in environmental 
peaks and effects on populations make negligible difference.

@@

\begin{figure}
\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseA.pdf}\\
\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseB.pdf}\\
\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v01.pdf}\\
\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v02.pdf}
\caption{Theoretical case studies A (panels a-d), B (panels e-h) and C (i-p, two alternative sets of 
parameters used for i-l and m-p, respectively). Left panels (a, e, i, m) show the terms on the right side
of (\ref{rhoww_T1to3}). On those panels, the green line is 
$\frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(1)} \epsilon^{(1)}}$, quantifying 
population synchrony due to the direct Moran effects of $\epsilon^{(1)}$. The blue line
is $\frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(2)} \epsilon^{(2)}}$, quantifying 
population synchrony due to the direct Moran effects of $\epsilon^{(2)}$.
The dashed green-blue line is 
$\frac{2 \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right)}{|f_{\mcB}|^2}$, 
which is population synchrony due to interacting Moran effects.
Middle panels show the constituent parts of the theory (labelled). For these complex functions 
of timescale, the magnitude is displayed on the vertical axis and the phase is displayed with color.
See text for all interpretations.
The functions plotted on d, h, l, p are those on a, e, i, m, respectively, times 
$|f_{\mcB}|^2$, thereby showing the influence of autoregressive population effects on
synchrony [see (\ref{rhoww_T1to3})].
Model parameters were: $N=5$ and $b_1=0.4$ for all case studies; 
$p_1^{(1)}=1.1$ and $p_0^{(2)}=0.8$ for case study A; 
$p_3^{(1)}=1.1$ and $p_0^{(2)}=0.8$ for case study B;
$p_1^{(1)}=3$, $p_0^{(2)}=1.5$ 
%$c_1=0$ and $c_2=-4/9$ 
for case study C. For C, peaks in the periodic noise process $\epsilon^{(2)}$
lagged peaks in the periodic process $\epsilon^{(1)}$ by 1 quarter
in panels i-l, and $\epsilon^{(1)}$ lagged $\epsilon^{(2)}$ by the same
amount in m-p.
See SI section \ref{SM-sect:examples_cross_sync} and Fig. \ref{SM-fig:rhosForC} for 
parameters associated with the noise for each case study,
and see Methods and SI section \ref{SM-sect:details_case_studies} for additional details.
Synch. contrib. = Synchrony contribution.}\label{fig:TheoryCaseA}
\end{figure}
